\clearpage

{\Large Parts of the thesis have appeared in the following publications.}
\begin{center}

\begin{itemize}
 \item \href{http://dl.acm.org/citation.cfm?id=2745955}{{\em Minimizing Coordination in Replicated Systems.}}\\
  Cheng Li, Jo{\~a}o Leit{\~ a}o, Allen Clement, Nuno Pregui{\c c}a, and Rodrigo Rodrigues.\\
  \emph{PaPoC'15}, Apr. 2015, Bordeaux, France
 \item \href{https://www.usenix.org/system/files/conference/atc14/atc14-paper-li_cheng.pdf}{{\em Automating the Choice of Consistency Levels in Replicated Systems.}}\\
  Cheng Li, Jo{\~a}o Leit{\~ a}o, Allen Clement, Nuno Pregui{\c c}a, Rodrigo Rodrigues and Viktor Vafeiadis.\\
 \emph{USENIX ATC'14}, Jun. 2014, Philadelphia, PA, USA
 \item \href{https://www.usenix.org/system/files/conference/osdi12/osdi12-final-162.pdf}{{\em Making Geo-Replicated Systems Fast as Possible, Consistent when Necessary.}}\\
  Cheng Li, Daniel Porto, Allen Clement, Rodrigo Rodrigues, Nuno Pregui{\c c}a, and Johannes Gehrke.\\
 \emph{OSDI'12}, Oct. 2012, Hollywood, CA, USA
 \item \href{http://www.mpi-sws.org/~chengli/trs/rbTR.pdf}{{\em Making Geo-Replicated Systems Fast as Possible, Consistent when Necessary.}}\\
Cheng Li, Daniel Porto, Allen Clement, Rodrigo Rodrigues, Nuno Pregui{\c c}a, and Johannes Gehrke.\\
 \emph{Technical Report}, MPI-SWS, 2012
\end{itemize}
\end{center}

\clearpage
{\Large Additional publications published while at MPI-SWS.}
\begin{center}
\begin{itemize}
 \item {\em WeaQL: Scaling Relational Databases through Weak Consistency.}\\
  David Lopes, Cheng Li, Jo{\~a}o Leit{\~ a}o, Nuno Pregui{\c c}a, and Rodrigo Rodrigues.\\
  \emph{Submitted to ICDCS'16}
 \item \href{http://dl.acm.org/citation.cfm?id=2741979}{{\em Visigoth Fault tolerance.}}\\
  Daniel Porto, Jo{\~a}o Leit{\~ a}o, Cheng Li, Allen Clement, Aniket Kate, Flavio Junqueira, and Rodrigo Rodrigues.\\
 \emph{EuroSys'15}, Apr. 2015, Bordeaux, France
  \item \href{http://www.mpi-sws.org/~dcfp/files/vftTR.pdf}{{\em  Lower Bound and Correctness Proofs for Consensus in the Visigoth Model.}}\\
  Daniel Porto, Jo{\~a}o Leit{\~ a}o, Cheng Li, Allen Clement, Aniket Kate, Flavio Junqueira, and Rodrigo Rodrigues.\\
  \emph{Technical Report}, Nova University of Lisbon, 2015
 \item \href{http://dl.acm.org/citation.cfm?id=1966465}{{\em Finding Complex Concurrency Bugs in Large Multi-Threaded Applications.}}\\
 Pedro Fonseca, Cheng Li, and Rodrigo Rodrigues.\\
 \emph{EuroSys'11}, Apr. 2011, Salzburg, Austria
 \item \href{http://www.mpi-sws.org/~rodrigo/dsn2010-concurrencybugs.pdf}{{\em A study of the Internal and External Effects of Concurrency Bugs.}}\\
 Pedro Fonseca, Cheng Li, Vishal Singhal and Rodrigo Rodrigues.\\
 \emph{DSN'10}, Jun. 2010, Chicago, USA
\end{itemize}
\end{center}


\if 0
\vspace*{\stretch{2}}
\begin{center}
{\Huge \textbf{\textsf{Building Fast and Consistent (Geo-)Replicated Systems: from Principles to Practice}}} \\[4em]

{\huge \textsf{Thesis for obtaining the title of Doctor of Engineering Science of the Faculty
of Natural Science and Technology of Saarland University}} \\[2em]%
\end{center}
\vspace*{\stretch{4}}
\begin{center}
{\LARGE From}\\[1em]
{\LARGE \textsf{Cheng Li}}
%\end{center}
%\vspace*{\stretch{4}}
%\begin{center}
%{\Large \textbf{\textsf{Thesis}}} \\
~\\
%\sf Version 1 \\
%~\\
%zur Erlangung des Grades \\
%des Doktors der Ingenieurwissenschaften \\
%der Naturwissenschaftlich-Technischen Fakult\"aten \\
%der Universit\"at des Saarlandes \\
%~\\
~\\
%Eingereicht Saarbr\"ucken, 17.~December 2015
Saarbr\"ucken
December 2015
\end{center}
\fi
