\section{Restriction inference}
\label{ch:por:sect:infer}
When replicating a service under \PRCN, the first step is to
infer restrictions to ensure two important system 
properties, namely, state convergence and invariant preservation. 
The major challenge we face is to identify a small set of restrictions for
making the replicated service eventually converge and never violate invariants so
that the amount of required coordination is minimal.
With regard to state convergence, we take a similar
methodology adopted in prior research~\cite{Shapiro2011CRDTs,
Li2012RedBlue, Li2014SIEVE}, which is to check operation commutativity.
However, unlike \RBCN, under which all operations that are not globally commutative must be totally ordered,
\PRCN\ only requires that an operation must be ordered w.r.t another one if
they do not commute.

To always preserve application-specific invariants,
instead of totally ordering all non-invariant safe
shadow operations, i.e., those that potentially transition from a valid state
to an invalid one, we try to isolate
the operations that exclusively contribute to an invariant violation from
the rest. To do so, we introduce
a new concept, called an {\tt I-conflict set}, which defines a minimal set of
shadow operations that lead to an invariant violation when they are running
concurrently in a coordination-free manner. By minimal, we mean that by removing 
any shadow operation from such a set, the violation will no longer persist. 
To identify a minimal set of restrictions, we first perform an analysis over any subsets of the shadow 
operation set to discover all {\tt I-conflict} sets. Then, for any such set, adding a 
restriction between any pair of its operations is sufficient to eliminate 
the problematic executions.

Next, we present the definitions, theorems, proofs and algorithms
regarding the restriction set identification and refinement.

\input{./por/convergence}
\input{./por/invariant}
